Raising the Role of Vocabulary Hubs for Semantic Data Interoperability in Dataspaces

### 2024-06-12 EBDVF-2024/paper.qmd
Added details for use cases and conclusion

# Projects and Use Cases
In the expanding data economy landscape supported by dataspaces, 
we are working on several projects which address practical data sharing needs and were we experience the importance of semantic interoperability to fulfill various use cases. 
Here we present 3 examples from the [DataBri-X](https://databri-x.eu/) and [UNDERPIN](https://underpinproject.eu/) projects that cover various application domains.

## Energy Dataspace
In this use case, we run simulations for energy communities, which predict the behavior of the energy grid. 
To provide precise predictions, there is the need of high amounts of example data. 
This data is of numeric type and different tabular data sources need to be semantically consolidated.

## Legal Dataspace
Analyzing documents in the legal domain is a highly challenging task because of the complexity of legal information. 
We provide recommender services for insights into corpora of legal documents
that identify the meaning via a legal knowledge graph, NLP analysis and semantic annotations.

@@

@@

# Conclusion and Future Work
We presented an approach for dynamically binding datasets to semantic descriptions in a dataspace's Vocabulary Hub,
therefore facilitating data harmonization and easier data consumption.
We have implemented a first prototype of our approach.

The next steps include:
- Bringing it into an actual dataspace in practice, representing and integrating datasets from different dataspace participants.
- Implementing use cases for richer discoverability and harmonized querying.
Also, the ML should actually benefit regarding quality in practice, which still needs to be verified.

>>>>

# Projects and Use Cases
In the expanding data economy landscape supported by dataspaces, 
we are working on several projects which address practical data sharing needs and were we experience the importance of semantic interoperability to fulfill various use cases. 
our main focus to develop our solution is the [UNDERPIN](https://underpinproject.eu/) project with use cases of refineries and wind farms,
but also other project, like [DataBri-X](https://databri-x.eu/) provide use cases which were considered.

@@

### The Refinery Dataspace
Refineries can be optimized regarding costs and efficiency by considering not only individual components,
but the production chain as a connected system.
In this use case, we aim to improve the maintenance process as well as the decision making for preventive maintenance
so as to minimize the downtime and the impact on the production capabilities.

### The Wind Farm Dataspace
Operating wind farms involves high costs for maintaining the wind turbines.
Predicting potential failures and optimizing the maintenance operations to minimize the costs can be business critical.
For establishing predictive maintenance, we need high amounts of high qualitative training data,
which can be achieved by consolidation different wind farm measurement sources.
This use case provides different kinds of tabular data, where consolidation needs semantic harmonization of data.

## Energy and Legal Dataspace
Beyond the UNDERPIN project as the main use case provider, we also consider use cases from other projects.
We present 2 examples from the DataBri-X project, which work on different content types of data and have different usage scenarios.
We considered these as well when designing our solution to make it applicable to a broader range of use cases.

### The Energy Dataspace
In this use case, we run simulations for energy communities, which predict the behavior of the energy grid. 
To provide precise predictions, there is the need of high amounts of example data. 
This data is of numeric type and different tabular data sources need to be semantically consolidated.

### The Legal Dataspace
Analyzing documents in the legal domain is a highly challenging task because of the complexity of legal information. 
We provide recommender services for insights into corpora of legal documents
that identify the meaning via a legal knowledge graph, NLP analysis and semantic annotations.

@@

# Conclusion and Future Work
We presented an approach for dynamically binding datasets to semantic descriptions in a dataspace's Vocabulary Hub,
therefore facilitating data harmonization and easier data consumption.
We have implemented a first prototype of our approach, which builds on the integration of GraphDB and PoolParty.

The next steps include:
- Bringing it into an actual dataspace in practice, representing and integrating datasets from different dataspace participants.
- Implementing use cases for richer discoverability, harmonized querying and support for different content types of structured and unstructured data.
Also, the ML should actually benefit regarding quality in practice, which still needs to be verified.
